---
title: "STAT 440 Statistical Data Management - Fall 2021"
output: html_document
---
## Week 09 Notes
### Created by Christopher Kinson


***


## Table of Contents

- [Summarizing data](#summarizing)
- [Combining Data](#combining)


***


## <a name="summarizing"></a>Summarizing data

Another important aspect of data wrangling is to summarize or aggregate data. This may also be considered as applying summary functions, such as the sum `sum()`, mean `mean()`, median `median()`, variance `var()`, or standard deviation `sd()`, to grouped data aka "group processing". Grouped data can be any data with a categorical variable or factor as a column. This task comes in handy when we want to know statistical values for each group. To accomplish summarization, sometimes we can leverage the way the data are arranged (or sorted). Other times, the arrangement has no bearing on our ability to aggregate. Ideally, we want the result to be a data frame, tibble, or other recursive object when possible.

Working with the City of Urbana's [Rental Inspection Grades Listing Data as tab-separated .txt - GHE](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/rental-inspections-grades-data03.txt) or the Box data URL https://uofi.box.com/shared/static/0j11not1cqmonbrwy0l8zmzr27fafm2j.txt, we may have to use loops and conditional execution to achieve the grouped summaries as a base R approach. I didn't mention it explicitly, but we've done this kind of work in Week06. For review, let's compute the proportion for each inspection letter grade.

```{r}
library(tidyverse)
RentalsData <- read_delim("https://uofi.box.com/shared/static/0j11not1cqmonbrwy0l8zmzr27fafm2j.txt", delim="\t")

g <- sort(unique(RentalsData$Grade))
groups <- rep(0,length(g))
for (i in 1:length(groups)){
 groups[i] <- nrow(RentalsData[RentalsData$Grade==g[i],"Grade"])
}
data.frame(grades = g, count_grade=groups, grade_proportion = groups/sum(groups), row.names = NULL)
```

In order to summarize group data with the tidyverse, we need two actions to work in tandem: `group_by()` and `summarise()`. 

**summarize() is the same thing as summarise() in R for all you Americans.**

The `group_by()` function allows us to do data processing separately for each group in the data frame. By itself, the function does not affect the data outwardly. The group processing takes place once we combine this function with other **dplyr** functions, especially `summarise()`. The `group_by()` can work with `mutate()` and `arrange()` as well when specified.

We can create grouped summaries of data with the `summarise` function. By itself it collapses a data frame into a vector by aggregating. **This implies that we first group the data, then summarise.** Also notice the difference between `mutate()` and `summarise()` after the `group_by()`.

```{r}
gg <- group_by(RentalsData, Grade)
gg
summarise(gg, count_grade = n())
summarise(group_by(RentalsData, Grade), count_grade = n(), proportion_grade=count_grade/nrow(RentalsData)) # same as above
```


***


## <a name="combining"></a>Combining data

What happens when you need to work with multiple datasets at once? What happens when the one dataset you have is not enough information? Where do you get the additional information? Combining data sets is a very useful data wrangling operation. Grabbing information from another dataset and adding it to your current one potentially increases your information. Combining data could mean different things in various disciplines or the same thing with different terms such as concatenating, merging, binding, appending, or joining.

In R, concatenating is the act of combining objects or strings together and is typically done with `c()` or `paste()` or `glue()` in **glue**. In order to use the `glue()` function, we may have to install the glue package separately and call it separately as in `glue::glue()`.

In R, binding (or appending) is the act of combining two or more objects by stacking one on top of the other such as `rbind()` in R (or `bind_rows()` in **dplyr**) or stacking one next to the other such as `cbind()` in base R (or `bind_cols()` in **dplyr**).

Merging (or joining) usually implies combining two or more objects with different columns of information into one single object. This merging would require each of the different data objects to have one column in common with a unique identifying information such as an ID variable or geographic location. There are at least 3 situations that can occur when merging objects. 

1. Observations in the two (or more) separate objects could not match each other.

**Data 1**  

ID | Salary
---|---
A | $10K
B | $11K
D | $12K

**Data 2**  

ID | Number
---|---
C | 2175551234
E | 2175551235
F | 2175551236

**Merged Data**  

ID | Salary | Number
---|---|---
A | $10K |
B | $11K |
D | $12K |
C |  | 2175551234
E |  | 2175551235
F |  | 2175551236

2. Observations in the two (or more) separate objects could match each other one-to-one.

**Data 1**  

ID | Salary
---|---
A | $10K
B | $11K
D | $12K

**Data 2**  

ID | Number
---|---
A | 2175551214
B | 2175551224
D | 2175551244

**Merged Data**  

ID | Salary | Number
---|---|---
A | $10K | 2175551214
B | $11K | 2175551224
D | $12K | 2175551244

3. Observations in the two (or more) separate objects could match each other one-to-many (or many-to-one).

**Data 1**  

ID | Salary
---|---
A | $10K
D | $12K

**Data 2** 

ID | Number
---|---
A | 2175551214
A | 2175551204
D | 2175551244

**Merged Data**  

ID | Salary | Number
---|---|---
A | $10K | 2175551214
A | $10K | 2175551204
D | $12K | 2175551244

How we merge (or join) the data depends on which of the three situations is intended for the data management. Only keeping the matches (#2 and #3 above) could be accomplished using an inner join (`inner_join()` in **dplyr** or `merge(...,all=FALSE)` in base R) . Keeping the matches (#2 and #3 above) and non-matches (#1 above) could be accomplished using a full join (`full_join()` in **dplyr** or `merge(...,all=TRUE)` in base R). Whenever the common column of the different data objects contain the same information but have different column names, the easiest fix is to rename the column in one of the two objects (`rename()` in **dplyr**).

Let's combine the owner addresses (scraped and saved from Week 03 Notes) with the `RentalsData` as [owners-addresses .csv - GHE](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/owners-addresses.csv) or the Box data URL https://uofi.box.com/shared/static/u6coxibtzx3mith23bzk4rysu923g160.csv. Doing this combining is quite simple because we have only one column and the same number of elements in the RentalsData and owners-addresses.

```{r}
owners_addresses <- read_csv("https://uofi.box.com/shared/static/u6coxibtzx3mith23bzk4rysu923g160.csv")

RentalsData2 <- bind_cols(RentalsData,owners_addresses)

slice_head(RentalsData2, n=10)
```

**SN: The `distinct()` function is a dplyr function that works much like base R's `unique()` function. It may come in handy when wanting to find the number of unique individuals of a particular subset of a dataset.**

```{r}
unique(RentalsData$Grade)

distinct(RentalsData,Grade)

distinct(RentalsData, Grade, .keep_all = TRUE)
```

**There are a host of other functions and operations to be discovered in tidyverse's dplyr. Please check out the RStudio cheatsheets for dplyr https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf.**

#### END OF NOTES