---
title: "prosperity_1125"
author: "Jonathan Kang (jk36)"
date: '2022-11-25'
output: html_document
---

```{r results='hide', message=FALSE, echo=FALSE}
library(dplyr)
library(tidyverse)
library(readxl)
library(sqldf)
library(ggplot2)
library(GGally)
library(reshape2)
library(MASS)
library(mclust)

```
# Import Data
```{r}
free_prosp_full_dataset <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    na = "no data")

free_prosp_data <- sqldf("SELECT Region as region,
              Country as country,
              `Freedom rank 2021` as freedom_rank,
              `Freedom category 2021` as freedom_category, 
              `Freedom score 2021` as freedom_score,
              `Economic Freedom score 2021` as economic_freedom_score,
              `Political Freedom score 2021` as political_freedom_score,
              `Legal Freedom score 2021` as legal_freedom_score,
              `Prosperity rank 2021` as prosperity_rank,
              `Prosperity category 2021` as prosperity_cateogry,
              `Prosperity score 2021` as prosperity_score,
              `Income score 2021` as income_score,
              `Environment score 2021` as environment_score,
              `Minority Rights  score 2021` as minority_rights_score,
              `Health score 2021` as health_score,
              `Happiness score 2021` as happiness_score
       FROM free_prosp_full_dataset;
      ")
glimpse(free_prosp_data)
```

No ties in ranking

# Exploratory Visualizations

## Overall Dataset
```{r}
## SCORES
### Apply log transformation to satisfy normality assumptions
model_data <- free_prosp_data %>%
  dplyr::select(region, freedom_score, prosperity_score) %>%
  mutate(log_prosperity = log(prosperity_score))
lin_model = lm(freedom_score ~ log_prosperity, model_data)

## Check normalities, outliers, heteroscedasticity
# plot(model_data$freedom_score, model_data$log_prosperity)
# plot(lin_model)
summary(lin_model)

ggplot(data = model_data, aes(x = (freedom_score), y = prosperity_score)) +
  geom_point(aes(color = region)) + 
  geom_smooth(method = "glm", se = T) + 
  labs(title = "Freedom Score v Prosperity Score") + 
  xlab("Freedom Score") +
  ylab("Prosperity Score")


```

Multiple R-squared:  0.6353



## Economic
### Overall
```{r}
econ_subs_dataset <- free_prosp_data %>%
  dplyr::select(region, economic_freedom_score, political_freedom_score, legal_freedom_score)
ggpairs(data=econ_subs_dataset,1, mapping=aes(color=region))
ggpairs(data=econ_subs_dataset,2:4, mapping=aes(color=region))


```

## Political

```{r, warning = FALSE}
polit_subs_dataset <- free_prosp_data %>%
  dplyr::select(region, legal_freedom_score, prosperity_score, income_score, environment_score, minority_rights_score, health_score,happiness_score)
# ggpairs(data=polit_subs_dataset, mapping=aes(color=region))
ggpairs(data=polit_subs_dataset,2:4, mapping=aes(color=region))
ggpairs(data=polit_subs_dataset,5:8, mapping=aes(color=region))

```

### Cluster the groups by using Mixture of Multivariate Gaussian Model 

```{r}
table(free_prosp_data$prosperity_cateogry, free_prosp_data$freedom_category)

prosp_counts = free_prosp_data %>% dplyr::select(prosperity_cateogry) %>% group_by(prosperity_cateogry) %>% count()
order = c( "Prosperous", "Mostly Prosperous", "Mostly Unprosperous", "Unprosperous")
prosp_counts <- prosp_counts %>%
  mutate(prosperity_cateogry = factor(prosperity_cateogry, levels = order)) %>%
  arrange(prosperity_cateogry)

ggplot(prosp_counts) +
  geom_col(aes(x = prosperity_cateogry, y = n))

freedom_counts = free_prosp_data %>% dplyr::select(freedom_category) %>% group_by(freedom_category) %>% count()
order = c("Free", "Mostly Free", "Mostly Unfree", "Unfree")
freedom_counts <- freedom_counts %>%
  mutate(freedom_category = factor(freedom_category, levels = order)) %>%
  arrange(freedom_category)

ggplot(freedom_counts) +
  geom_col(aes(x = freedom_category, y = n))
```

Results show that there is a possibility we can cluster into 2 groups by comparing the scores.  
We will proceed with an unsupervised learning method of using Gaussian Mixture Models:

```{r}
colors = c("Freedom" = "Red", "Prosperity" = "Blue")
ggplot(free_prosp_data) +
  geom_density(aes(x = freedom_score, color="Freedom")) + 
  geom_density(aes(x = prosperity_score, color="Prosperity")) +
  xlab("Score") + 
  labs(title = "Density plot for Prosperity and Freedom Scores", color = "Legend") +
  scale_color_manual(values = colors)

# gauss_data <- free_prosp_data %>% select(freedom_score, prosperity_score)
# plot(mclustBIC(model_data[c(2,3)]))
gaussian_mix_model = densityMclust(data=model_data[c(2,3)], G=2)
# summary(gaussian_mix_model)
plot(gaussian_mix_model, what = "density", type = "hdr", data = model_data[c(2,3)], points.cex = 0.5) +
title("Mixture Gaussian Clusters Between Freedom and Prosperity")
```


# Correlation Plots for Each Category
Here, we will look at various correlation plots to determine the most important variable by comparing the correlation coefficients.

## Economic Freedom
```{r warning=FALSE}
economic_freedom_scores <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    na = "no data")[c(1,3,7)]
# Property Rights
prop_rights_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
  sheet = "Property Rights time")
prop_rights <- prop_rights_full[c(1,4)]
# Trade Freedom
trade_free_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
  sheet = "Trade Freedom time")
trade_free <- trade_free_full[c(1,4)]
# Invesetment Freedom
invest_free_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
  sheet = "Investment Freedom time")
invest_free <- invest_free_full[c(1,4)]
# Women's Economic Freedom
women_econ_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
  sheet = "Women's Economic freedom time")
women_econ <- women_econ_full[c(1,4)]

economic_freedom <- economic_freedom_scores %>%
  left_join(prop_rights, by="Country") %>%
  left_join(trade_free, by="Country") %>%
  left_join(invest_free, by="Country") %>%
  left_join(women_econ, by="Country")

ggpairs(economic_freedom[c(3:7)])


```

## Political Freedom
```{r}
political_freedom_scores <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    na = "no data")[c(1,3,8)]

# Constraints on Government
const_on_govt_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Constraints on Government time")
const_on_govt <- const_on_govt_full[c(1,4)]
# Political Rights
political_rights_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Political Rights time")
political_rights <- political_rights_full[c(1,4)]
# Civil Liberties
civil_liberties_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Civil Liberties time")
civil_liberties <- civil_liberties_full[c(1,4)]

political_freedom <- political_freedom_scores %>%
  left_join(const_on_govt, by="Country") %>%
  left_join(political_rights, by="Country") %>%
  left_join(civil_liberties, by="Country")

ggpairs(political_freedom,3:6)
```


## Legal Freedom

```{r warning=FALSE}
legal_freedom_scores <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    na = "no data")[c(1,3,9)]

# Judicial Effectiveness
judicial_effect_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Judicial Effectiveness time")
judicial_effect <- judicial_effect_full[c(1,4)]
# Efficient Judiciary
eff_jud_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Efficient Judiciary time")
eff_jud <- eff_jud_full[c(1,4)]
# Civil Justice
civil_justice_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Civil Justice time")
civil_justice <- civil_justice_full[c(1,4)]
# Criminal Justice
criminal_justice_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Criminal Justice time")
criminal_justice <- criminal_justice_full[c(1,4)]
# Government Integrity
govt_integrity_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = " Government Integrity time")
govt_integrity <- govt_integrity_full[c(1,4)]
# Public Disclosure
public_disc_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Public Disclosure time")
public_disc <- public_disc_full[c(1,4)]
# State Capacity
state_cap_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "State Capacity time")
state_cap <- state_cap_full[c(1,4)]
# Order and Security
order_security_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Order and Security time")
order_security <- order_security_full[c(1,4)]
# Regulatory Effectivness
reg_effectiveness_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Regulatory Effectivness time")
reg_effectiveness <- reg_effectiveness_full[c(1,4)]

legal_freedom <- legal_freedom_scores %>%
  left_join(judicial_effect, by="Country") %>%
  left_join(eff_jud, by="Country") %>%
  left_join(civil_justice, by="Country") %>%
  left_join(criminal_justice, by="Country") %>%
  left_join(govt_integrity, by="Country") %>%
  left_join(public_disc, by="Country") %>%
  left_join(state_cap, by="Country") %>%
  left_join(order_security, by="Country") %>%
  left_join(reg_effectiveness, by="Country")

ggpairs(legal_freedom,3:12)
```

## Prosperity
```{r}
prosperity_scores <- (read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    na = "no data")[c(1,3,12:17)])

# ggpairs(prosperity_scores,3:8)
```

## Overall Freedom Scores
```{r}
categories <- free_prosp_full_dataset[c(1,3,6:9,12:17)]
# ggpairs(categories,3:6)
```



# Determine countries with largest changes over 15 years
## Freedom
```{r}
freedom_changes_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = " Freedom Index time")

freedom_changes <- freedom_changes_full %>%
  mutate(difference = `Freedom Score 2021` - `Freedom Score 2006`) %>%
  mutate(abs_percentages = abs(`Freedom difference percentage over sample`)) %>%
  dplyr::select(Country, Region, abs_percentages, difference)

freedom_perc <- freedom_changes %>%
  arrange(desc(abs_percentages))
freedom_perc[!duplicated(freedom_perc$Region),]

freedom_diff <- freedom_changes %>%
  arrange(desc(difference))
freedom_diff[!duplicated(freedom_diff$Region),]
```
## Prosperity
```{r}
prosperity_changes_full <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = " Prosperity Index time")

prosperity_changes <- prosperity_changes_full %>%
  mutate(percentages = (`Prosperity Index 2021` - `Prosperity Index 2006`) / `Prosperity Index 2006`) %>%
  mutate(abs_percentages = abs(percentages) * 100) %>%
  dplyr::select(Country, Region, abs_percentages, `Difference over sample`)

prosperity_perc <- prosperity_changes %>%
  arrange(desc(abs_percentages)) 
prosperity_perc[!duplicated(prosperity_perc$Region),]

prosperity_diff <- prosperity_changes %>%
  arrange(desc(`Difference over sample`))
prosperity_diff[!duplicated(prosperity_diff$Region),]

```


## Categories and their highest correlating factor
Economic Freedom <- Property Rights
Political Freedom <- Civil Liberties (or Political Rights)
Legal Freedom <- Regulatory Effectiveness

Prosperity <- Environment
Freedom <- Political Freedom

## Countries with largest changes 
```{r Freedom}
freedom_perc[!duplicated(freedom_perc$Region),]
```
```{r Prosperity}
prosperity_perc[!duplicated(prosperity_perc$Region),]
```

## Linear Regression for Important Variables

```{r}
# Economic Freedom <- Property Rights
econ_free_time <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Economic Freedom time")[c(1,3,4:7)]
prop_rights_time <- prop_rights_full[c(1,4:7)]
  #pivot_longer(!Country)
econ_free_temp <- econ_free_time %>% left_join(prop_rights_time, by="Country") %>% na.omit()
ggplot(data=econ_free_temp, aes(x = `Economic Freedom score 2021`, y=`Property Rights score 2021`)) +
  geom_point(aes(color = Region)) + geom_smooth(method = "glm") + labs(title = "Economic Freedom v Property 2021")

# free_prosp_full_dataset

# Political Freedom <- Civil Liberties (or Political Rights)
polit_index_time <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Political Freedom time")[c(1,3,4:7)]

civil_lib_time <- civil_liberties_full[c(1,4:7)]

polit_free_time <- polit_index_time %>% left_join(civil_lib_time, by="Country")
ggplot(data=polit_free_time, aes(x = `Political Freedom score 2021`, y=`Civil liberties score 2021`)) +
  geom_point(aes(color = Region)) + geom_smooth(method = "glm") + labs(title = "Political Freedom v Civil Liberties 2021")

# Legal Freedom <- Regulatory Effectiveness
legal_index_time <- read_excel("Freedom-and-Prosperity-Indexes-Full-Data-Set.xlsx", 
    sheet = "Legal Freedom time")[c(1,3,4:7)]

reg_eff_time <- reg_effectiveness_full[c(1,4:7)]

legal_free_time <- legal_index_time %>% left_join(reg_eff_time, by="Country") %>% na.omit
ggplot(data=legal_free_time, aes(x = `Legal Freedom score 2021`, y=`Regulatory Effectivness score 2021`)) +
  geom_point(aes(color = Region)) + geom_smooth(method = "glm") + labs(title = "Legal Freedom v Regulatory Effectiveness 2021")


```


```{r}
#test
free_prosp_data %>% filter(country == "Hungary")
freedom_perc
freedom_perc %>%
  filter(Country == "United States of America")

prosperity_perc %>%
  filter(Country == "United States of America")

freedom_perc %>% dplyr::select(-Country) %>% group_by(Region) %>% summarise(mean_p = mean(abs_percentages), mean_d = mean(difference)) %>%
  filter(Region == "WE/EU")


```


