---
title: "CMI Data exploratory"
author: "Jonathan Kang (jk36)"
date: '2022-09-23'
output:
  pdf_document: default
  html_document: default
---

```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(readr)
library(sqldf)
library(lubridate)
library(ggplot2)
library(plotly)
library(weathermetrics)
library(ggridges)
```

### Import .txt data files

```{r}
# original
ogCMIDAY <- read.delim("~/Documents/College/FA22/URES/RawData/CMIDAY.txt")
units <- head(ogCMIDAY,1) # units for each category

# using bash, remove the second row with the units, create new file without the units
  # sed 2d CMIDAY.txt > CMIDAY_rmunits.txt
CMIDAY_rmunits <- read.delim("~/Documents/College/FA22/URES/RawData/CMIDAY_rmunits.txt")
# glimpse(CMIDAY_rmunits)

CMIDAY_data <- head(CMIDAY_rmunits, -10) # excess notes at bottom
# head(CMIDAY_data)
```

Notes from txt (removed in CMIDAY_data): M = Missing data;\
E = Estimated data;\
suggested cication = Water and Atmospheric Resources Monitoring Program. Illinois Climate Network. (2022). Illinois State Water Survey, 2204 Griffith Drive, Champaign, IL 61820-7495. <http://dx.doi.org/10.13012/J8MW2F2Q>.

## Modify Data

NAs are fixed by finding the mean of 7 days before and after.

```{r, warning=FALSE}
data <- sqldf("
SELECT year, month, day, avg_rel_hum, avg_air_temp as avg_air_temp_f
FROM CMIDAY_data
")

data$avg_rel_hum <- as.numeric(data$avg_rel_hum)
data$avg_air_temp_f <- as.numeric(data$avg_air_temp_f)

## Review NA Rows
# rows = grep(T, is.na(data$avg_air_temp_f))
# slice(data, rows)
# data[rows,]
# rows = grep(T, is.na(data$avg_rel_hum))
# slice(data, rows)
# data[rows,]
#

# Fix the NA issues, then proceed with Celsius version and create time series for that
data2 = data
# for row 8326,8327
mean_temp1 = mean(c(data$avg_air_temp_f[(8325-6):8325],data$avg_air_temp_f[8328:(8328+6)]))
mean_hum1 = mean(c(data$avg_rel_hum[(8325-6):8325],data$avg_rel_hum[8328:(8328+6)]))
# for row 9213
mean_temp2 = mean(c(data$avg_air_temp_f[(9212-6):9212],data$avg_air_temp_f[9214:(9214+6)]))
mean_hum2 = mean(c(data$avg_rel_hum[(9212-6):9212],data$avg_rel_hum[9214:(9214+6)]))
# for row 9991,9992,9993
mean_temp3 = mean(c(data$avg_air_temp_f[(9990-6):9990],data$avg_air_temp_f[9994:(9994+6)]))
mean_hum3 = mean(c(data$avg_rel_hum[(9990-6):9990],data$avg_rel_hum[9994:(9994+6)]))
data2$avg_air_temp_f[8326:8327] = round(mean_temp1,3)
data2$avg_air_temp_f[9213] = round(mean_temp2,3)
data2$avg_air_temp_f[9991:9993] = round(mean_temp3,3)
data2$avg_rel_hum[8326:8327] = round(mean_hum1,3)
data2$avg_rel_hum[9213] = round(mean_hum2,3)
data2$avg_rel_hum[9991:9993] = round(mean_hum3,3)
data2 = data2[-11763,] # remove empty row

# # check nas
# rows = grep(T, is.na(data2$avg_air_temp_f))
# slice(data2, rows)
# data2[rows,]
# rows = grep(T, is.na(data2$avg_rel_hum))
# slice(data2, rows)
# data2[rows,]
##

data_Xna <- data2 %>% mutate(
  avg_air_temp_c = (avg_air_temp_f - 32) * (5/9),
  date_string = paste(year, month, day, sep="-"),
  date = as.Date(date_string)
)

glimpse(data_Xna)
```

### Time series (temp)

```{r, echo=FALSE}
time_series_faren <- plot_ly(data_Xna, type = 'scatter', mode = 'lines') %>%
  add_trace(x = ~date, y = ~avg_air_temp_f, line = list(color = "grey")) %>%
  layout(showlegend = F, title = "Average Air Temperature in Champaign (Farenheit)",
         yaxis = list(title="Degrees Farenheit"))
time_series_faren

time_series_cels <- plot_ly(data_Xna, type = 'scatter', mode = 'lines') %>%
  add_trace(x = ~date, y = ~avg_air_temp_c, line = list(color = "grey")) %>%
  layout(showlegend = F, title = "Average Air Temperature in Champaign (Celsius)",
         yaxis = list(title="Degrees Celsius"))
# time_series_cels

ggplot(data = data_Xna, aes(x = date, y = avg_air_temp_f)) +
  geom_line(color = "dark orange") + 
  labs(title = "Average Air Temperature in Champaign") +
  ylab("Average Temperature (F)") +
  xlab("Date")

```

## Heat Index

```{r}
data_HI <- sqldf("
SELECT date, avg_air_temp_f, avg_rel_hum, month
FROM data_Xna
")

# Bound of avg_rel_hum = [0,100] (only have vals > 100%)
large_rh = grep(T, data_HI$avg_rel_hum > 100)
# slice(data_HI, large_rh)
data_HI$avg_rel_hum[large_rh] = 100 

data_HI$heat_index <- heat.index(t = data_HI$avg_air_temp_f,
                                 rh = data_HI$avg_rel_hum,
                                 temperature.metric = "fahrenheit")

# the 85th percentile
heat_index_calc <- data_HI %>%
  filter(month == "7" | month == "8") %>%
  filter(date >= "1990-01-01" & date <"2021-01-01")

HI_index = quantile(heat_index_calc$heat_index, 0.85)
HI_index
```

The local 85th percentile = 83

```{r, echo=FALSE}
## Calculate the heat index (RH-adjusted temperature) time series for the two locations

time_series_RH <- ggplot(data_HI, aes(x=date, y=avg_rel_hum)) +
  geom_line(color="blue", size=0.5) +
  ggtitle("Relative Humidity") 
time_series_HI <- ggplot(data_HI, aes(x=date, y=heat_index)) +
  geom_line(color="orange",size=0.5) +
  geom_hline(yintercept=HI_index, linetype="dashed", color = "black", size=0.5) +
  geom_text(aes(as.Date("1989-02-16"), HI_index,label = HI_index, vjust = -1, hjust = 1)) +
  ggtitle("Heat Index")

# time_series_RH 
# time_series_HI
datar <- data_Xna %>%
  filter(year == 2010) %>%
  arrange(desc(month))
datar$month <- ordered(datar$month)  
ggplot(datar, aes(x = avg_air_temp_f, y = month)) +
  geom_density_ridges_gradient() + theme_ridges() +
  labs(title = 'Temperature of 2010 (Champaign)') + 
  xlab("Average Temperature (Farenheit)")

# interactive version
time_series_HI_2 <- plot_ly(data_HI, type = 'scatter', mode = 'lines') %>%
  add_trace(x = ~date, y = ~heat_index, line = list(color = "Orange")) %>%
  add_trace(x = ~date, y = ~HI_index, line = list(shape = "linear", color = "black", dash = 'dash')) %>%
  layout(showlegend = F, title = "Heat Index Interactive Time Series",
         yaxis = list(title="HI"))
time_series_HI_2

```

## Ridge Regression plot
```{r}
hi_ridge <- data_HI %>%
  # filter(year(date) == 2010) %>%
  arrange(date) %>%
  mutate(month = month(date))
hi_ridge$month <- ordered(hi_ridge$month)  
ggplot(hi_ridge) +
  geom_density_ridges_gradient( aes(x = heat_index, y = month)) + theme_ridges() +
  geom_vline(xintercept = HI_index) +
  labs(title = 'Heat Index (Champaign)') + 
  xlab("Heat Index")
```


# Variables

Determine the occurrences of heat wave:  
By definition, a heat wave is declared when HI > 85th percentile for 2 or more consecutive days.

```{r}
data_HW_setup <- data_HI %>%
  mutate(heat_index_85p = heat_index >= HI_index) %>% # heat_index_85p = T if heat_index > 85th
  select(date, heat_index, heat_index_85p) %>%
  filter(month(date) < 11 & month(date) > 4) %>% # within the hottest 6 months for easier look
  filter(heat_index_85p == T)

dates = data_HW_setup$date
sect = rep(-1,length(dates))
sect_index = 0
sect[1] = sect_index
for (day_index in (2:(length(dates)))) {
  curr = dates[day_index]
  yest = dates[day_index - 1] 
  tmrw = dates[day_index + 1]
  
  if(curr == (yest + 1)) { # if consec, add to section
    sect[day_index] = sect_index
  }
  else if(curr == (tmrw - 1)) { # if new consecutive, create and add new section
    sect_index = sect_index + 1
    sect[day_index] = sect_index
  }
}

data_HW_setup$section <- sect
data_HW <- data_HW_setup %>% filter(section >= 0) %>%
  select(date, heat_index, section)

head(data_HW,8)
```

### Pulses

Number of Heat Waves per Year.

```{r, echo=FALSE}
# count number of days for each heat wave occurrence
pulse_hw <- data_HW %>% 
  # count(section)
  group_by(year = year(date)) %>%
  count(section) 

# count number of heat waves for each year
pulse_hw_counts <- pulse_hw %>%
  group_by(year) %>%
  count(year)

# head(pulse_hw_counts)

pulse_plot <- ggplot(data = pulse_hw_counts, aes(x = year, y = n)) +
  geom_bar(stat="identity", fill="steelblue") +
  
  geom_hline(yintercept=mean(pulse_hw_counts$n), linetype="dashed", color = "black", size=0.2) +
  # geom_text(aes(x=pulse_hw_counts$year[1],y=mean(pulse_hw_counts$n),label = mean(pulse_hw_counts$n),2), vjust = -1, hjust = 1)) +
  
  labs(title="Pulses", x="Year", y = "Counts")
```

### Intensity

Mean excess heat index above threshold for each year.

```{r, echo=FALSE}
intensity_hw <- data_HW %>%
  group_by(year = year(date)) %>%
  summarize(mean_intensity = mean(heat_index - HI_index))

# head(intensity_hw)

intensity_plot <- ggplot(data = intensity_hw, aes(x = year, y = mean_intensity)) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_hline(yintercept=mean(intensity_hw$mean_intensity), linetype="dashed", color = "black", size=0.2) +
  labs(title="Mean Intensity (Excess Heat Index)", x="Year", y = "Intensity")
```

### Duration

Mean duration of a heat wave per year.

```{r, echo=FALSE}
colnames(pulse_hw) = c("year","section","days")
duration_hw <- pulse_hw %>%
  group_by(year) %>%
  summarise(mean_duration = mean(days))

# head(duration_hw)
duration_plot <- ggplot(data = duration_hw, aes(x = year, y = mean_duration)) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_hline(yintercept=mean(duration_hw$mean_duration), linetype="dashed", color = "black", size=0.2) +
  labs(title="Mean Duration", x="Year", y = "Duration")
```

### Season Length

The number of days between the start and end of a heat wave.

```{r, echo=FALSE}
season_hw <- data_HW %>%
  group_by(year = year(date)) %>%
  summarise(season_start = min(date), season_end = max(date),
            season_length = season_end - season_start + 1)
# head(season_hw)

season_plot <- ggplot(data = season_hw, aes(x = year, y = season_length)) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_hline(yintercept=mean(season_hw$season_length), linetype="dashed", color = "black", size=0.2) +
  labs(title="Season Length", x="Year", y = "Days")
```


### Joined Information

```{r}
season_hw2 <- season_hw %>%
  select(year, season_length)

important_factors <- pulse_hw_counts %>%
  rename("pulses" = "n") %>%
  left_join(intensity_hw, by = "year") %>%
  left_join(duration_hw, by = "year") %>%
  left_join(season_hw2, by = "year")

important_factors

# unique(important_factors$year)
par(mfrow = c(2,2))
# pulse_plot
# intensity_plot
# duration_plot
# season_plot

library(gridExtra)
grid.arrange(pulse_plot,intensity_plot, duration_plot, season_plot, nrow=2)




## Finding initial params for nleqslv
# lambda = E[pulses]
# mu = 1/E[intensity]
# theta = 1/E[duration]
mean(important_factors$pulses) 
1/mean(important_factors$mean_intensity) # inverse for initial ests 
1/mean(important_factors$mean_duration) # inverse for initial ests
```

Unusual Findings:  
- No heatwaves in 2004, 2008  
- Though 1997 only had 1 heat wave of 2 days, it was also the most intense year  



```{r}
group = NA
group[important_factors$year < 2000] = 1 #red
group[important_factors$year >= 2000 & important_factors$year < 2010] = 2 #blue
group[important_factors$year >= 2010] = 3 #purple

pairs(important_factors[2:5], pch = 18, col = c("red", "steelblue", "purple")[group])
```

TODO: Read article, check vars and corr for each variable
















