# STAT 440 Statistical Data Management - Fall 2021
## Week13 Lab03 Wed
### Only the driver should submit this assignment into their individual student repo. Save the filename as lab03-week13-wed-netID.Rmd (or .ipynb for Python users). Then render this file to .html when completed. 
#### Both .Rmd/.ipynb and .html files must be submitted by 11:00 am. Any lab submitted after 11:00 am will be deducted 1 point for being late.

**#1** Using Markdown syntax, make a non-numbered list with the first and last names of everyone at your station. Your name should appear in all caps.

Answers will vary, but must be a proper non-numbered list with Driver's name in all caps.

**#2** statement: Python allows users to create their own functions using the **def** keyword.

If the text in bold is the term that makes the statement true, then write TRUE below in all caps. If the text in bold is the term that makes the statement false, then write FALSE below in all caps and write the correct term in bold text in a new line under FALSE.

TRUE

**#3** **Import the following dataset using a programming language and its corresponding data URL. Ensure any date or time column(s) get properly formatted. Now using SQL, rename all columns such that there is no whitespace in the column names, arrange the Inspection ID column in ascending order, and print the last 5 rows of the resulting data using SQL.** Read the data descriptions to help verify the import was successful.

- Chicago Food Inspections Data Part 2 and its GHE data URL https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/chicago-food-inspections-data03.csv or Box data URL https://uofi.box.com/shared/static/zyn1kls0xxpjl3ld0m9w6cl2s3343gov.csv

- The dataset (.csv file) contains 9375 observations and 9 columns. The observations are places that serve food including grocery stores, butchers, bakeries, restaurants, school cafeterias, gas stations, and delis throughout the city limits of Chicago. These establishments pass, fail, or have certain conditions associated with passing the inspection. Read the two links - [link01](https://data.cityofchicago.org/api/assets/BAD5301B-681A-4202-9D25-51B2CAE672FF) and [link02](http://dev.cityofchicago.org/open%20data/data%20portal/2018/06/29/food-violations-changes.html) for more details about the data values. The original source is the City of Chicago Data Portal. **If you have difficulty accessing the two Chicago Food Inspections links from the original sources, please use these  [alt-link01](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/chicago-food-data-key01.pdf) and [alt-link02](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/chicago-food-data-key02.pdf) links instead.**

```{r}
library(tidyverse)
library(sqldf)

chicago_food_data02 <- read_delim("https://uofi.box.com/shared/static/zyn1kls0xxpjl3ld0m9w6cl2s3343gov.csv", col_types = cols(`Inspection Date` = col_date(format = "%m/%d/%Y")), delim="\t")

chicago_food_data02$`Inspection Date` <- as.character(chicago_food_data02$`Inspection Date`)

chicago_food_data <- sqldf("SELECT `Inspection ID` AS InspectionID, State, Zip, `Inspection Date` AS InspectionDate, `Inspection Type` AS InspectionType, Results, Violations, Latitude, Longitude
      FROM chicago_food_data02
      ORDER BY InspectionID ASC")

sqldf("SELECT *
      FROM (SELECT *
      FROM chicago_food_data
      ORDER BY InspectionID DESC
      LIMIT 5)
      ORDER BY InspectionID")
```

**#4** Create a new function called `string_summary` that takes 1 input and returns 1 output such that the following is honored:

- the input is 1 vector of string format
- the output is a data frame or tibble with 4 columns: 
  - column 1 named "words" : separates each word into a new row (words have only letters; not numbers; this column must be a unique set of words; no duplicates)
  - column 2 named "lengths": length of each word in "words"
  - column 3 named "vowels": number of vowels in each word in "words"
  - column 4 named "consonants": number of consonants in each word in "words"
- test your new function with the first row of the Violations column of the imported data in Problem 3.

**When I originally created this problem, I did not arrange the data in Problem 3 above, which is why the first row of the data has a legitimate string for Violations. This is not true after arranging the data by InspectionID.**

```{r}
string_summary <- function(input){
  newvec <- unique(str_extract_all(str_remove_all(tolower(input), "\\d"), boundary("word"))[[1]])
  tibble(words=newvec) %>%
    mutate(lengths=str_count(words), vowels=str_count(words, "a|e|i|o|u"), consonants=lengths-vowels)
}

string_summary(chicago_food_data$Violations[1])
string_summary(chicago_food_data$Violations[2])
string_summary(chicago_food_data$Violations[3]) #first two rows have NAs for Violations column
```

**#5** For a random sample of 20 rows, how many violations are recorded for each business? **You must answer in Markdown syntax in a complete sentence and show code as evidence of your answer.**

```{r}
sids <- sample(nrow(chicago_food_data),20)
tibble(violations=chicago_food_data$Violations[sids]) %>%
  mutate(violations_n=str_count(violations, "\\|")+1)
```
